---
title: "About"
author: "Chase Bookin"
date: "February 27, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
The Major League Baseball draft is an annual event in which the 30 MLB teams take turns selecting top high school and collegiate prospects. Once drafted, a player will choose to sign or, if they are a high school senior or college player with remaining eligibility, not sign and potentially get drafted again down the road.   

For top high school players especially, the magnitude of this decision is immense. While certain young players will sign for small contracts, eager to begin their professional career, others will forgo multi-million dollar signing bonuses and attend college. For some, the decision pays off handsomely and the prospect signs for more money as a polished college star. For others, they get injured or fail to perform in college and never play professional baseball.  

Money is obviously not the only factor in a high school prospect's decision - the team, prospect's health, and family influences among other things can play a significant role. However, given the amount of money and the opportunity many high school players opt to delay, it begs the question of how the decision plays out for these athletes.  

Do high school players who sign out of high school have a better chance of making it to the Major Leagues relative to their peers who attend college? What is the relationship between a player's decision of whether to sign and the longevity and production of their professional baseball career? Of drafted high school players that make it to the big leagues, do players that choose to attend college develop and experience more fruitful careers than players that sign immediately out of high school?  

The Major League Baseball draft is full of fascinating topics to analyze, and in this project I will use data to quantify high school players' career outcomes in relation to this decision.



```{r}
download.file("https://www.baseball-reference.com/data/war_daily_bat.txt", dest ="war.csv", mode = "wb")
```

Data: So far, I've loaded in a csv from BaseballReference containing WAR data for all players which essentially measures their production in the major leagues. It will be a good source for analyaing the final question I posed above. In addition, I have found draft data that may be useful; however, it seems tricky to download / scrape. I have been reading about scraping and should be able to get the data downloaded, joined, and cleaned soon. I will probably meet with Kaneesha sometime soon to discuss scraping, as it has been challenging to understand and try to this point. My GitHub repo is https://github.com/ChaseBook/final-project.git.





