---
title: "redraft backup"
author: "Chase Bookin"
date: "April 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(ggplot2)
library(janitor)
library(dplyr)
library(gt)
library(readxl)
library(sf)
library(scales)
library(magrittr)
library(haven)
library(infer)
library(Lahman)
library(xml2)
library(rvest)
library(DataCombine)
```


```{r, include=FALSE}
war <- read_csv("data/war.csv")
```


```{r, include=FALSE}

draft <- read_csv("data/draft_register.csv") %>%
  clean_names()

# grouping war data by player and summing career games played and war

war_joined <- war %>%
  group_by(mlb_ID) %>%
  summarize(g = sum(G),
            war = sum(as.numeric(WAR)))

```

```{r, include=FALSE}

# joined draft and war data by player mlb ID, then mutated teamname to condense
# the Angel's team names into a single team. I then arranged the data by year,
# grouped by player_id, and mutated a column called draft_mumber based on the
# row number (or occurence of that player's id) in the data. I did this because
# I wanted a column that tells for each draft a player is selected in, that it
# is the nth time that player has been selected. For example, if a player turns
# down the draft out of high school, after a year of junior college, and gets
# drafted after his junior year the draft_numb column should say 1,2,3 for each
# respective draft.

master <- draft %>%
  left_join(war_joined, by = c("mlbid" = "mlb_ID")) %>%
  mutate(teamname = as.factor(case_when(teamname %in% c("California Angels",
                                                        "Anahaim Angels",
                                                        "Los Angeles Angels of Anaheim") ~
                                                          "Los Angeles Angels",
                                        TRUE ~ teamname))) %>%
  arrange(year) %>%
  group_by(player_id) %>%
  mutate(draft_numb = row_number()) %>%
  ungroup()

# deleting useless columns

master$mlbid_1 <- NULL
master$x27 <- NULL

```

```{r}

# I made a new dataset where I filtered the master set for only unsigned
# highschool picks. I only selected the player_id column. I wanted to do this so
# that I could have a list of IDs and then rejoin these player IDs using a
# left_join to keep only the data from players who did not sign out of high
# school, but all years of their draft data rather than just the draft they
# turned down out of high school. That is what hs_unsigned is.

unsigned_ids <- master %>%
  filter(source == "H" & signed == "N") %>%
  select(player_id)

hs_unsigned <- unsigned_ids %>%
  left_join(master, by = "player_id") %>%
  arrange(player_id, desc(year))
```


```{r}

# filtering the unsigned high school data for only draft picks after 1990

hs_unsigned_trimyear <- hs_unsigned %>%
  filter(year >= 1990)

# filtered the data of unsigned high school picks with their subsequent drafts
# for only the first instance of the player being drafted, which is the draft
# afte their senior year of high school. For count_redrafted_hs, I filtered for
# draft_numb equal to 2 to see how many players were redrafted after originally
# turning the draft down.

count_unsigned_hs <- hs_unsigned_trimyear %>%
  filter(draft_numb == 1 & year <= 2015) %>%
  count() %>%
  pull(n)

# I create the ID table and re join it in order to get rid of data for redrafted
# players whose first draft out of high school was before the trim year (1990).

trimyear_ids <- hs_unsigned_trimyear %>%
  filter(draft_numb == 1 & year <= 2015) %>%
  select(player_id)

count_redrafted_hs <- trimyear_ids %>%
  left_join(hs_unsigned_trimyear, by = "player_id") %>%
  filter(draft_numb == 2) %>%
  count() %>%
  pull(n)

percent_redrafted_hs <- round(count_redrafted_hs / count_unsigned_hs * 100, 1)


# below, I repeat the same process for high school players drafted in the top 20
# rounds. I started by filtering the unsigned high school full dataset for only
# the initial high school draft and only those high school players taken in the
# first 20 rounds. I also filtered out players who were drafted out of high
# school after 2015, as it is still possible they are playing in college and may
# be redrafted later. I then re-joined the data to their subsequent drafts and
# found the percent of top 20 round high schoolers who are drafted again. AGAIN,
# only using data from players hwo were initially taken between 1990 and 2015.

# I filtered year less than 2015 to make sure we are selecting only top 20 round
# high school picks who turned down the draft between 1990 and 2015. Their
# subsequent drafts can occur any other year.

unsigned_hs_20_ids <- hs_unsigned_trimyear %>%
  filter(draft_numb == 1 & draft_round <= 20 & year <= 2015) %>%
  select(player_id)

hs_unsigned_20 <- unsigned_hs_20_ids %>%
  left_join(hs_unsigned_trimyear, by = "player_id")


first_draft_unsigned_20 <- hs_unsigned_20 %>%
  filter(draft_numb == 1)

avg_round_20_hs <- first_draft_unsigned_20 %>%
  summarize(mean_round = round(mean(draft_round), 2))%>%
  pull(mean_round)

count_unsigned_20 <- first_draft_unsigned_20 %>%
  count() %>%
  pull(n)

redrafted_20 <- hs_unsigned_20 %>%
  filter(draft_numb == 2)
  
avg_round_redrafted_20 <- redrafted_20 %>%
  summarize(mean_round = round(mean(draft_round),2)) %>%
  pull(mean_round)

count_redrafted_20 <- redrafted_20 %>%
  count() %>%
  pull(n)

percent_redrafted_20 <- round(count_redrafted_20 / count_unsigned_20 * 100, 1)

```
Findings from master redraft function (can be found in redraft_backup.Rmd) :
50.8 percent of players that turned down the draft out of high school from
1990-2015 were drafted again. 73.3 percent of players that turned down a top 20
round draft pick out of high school from 1990-2015 were drafted again. The
average initial round of top 20 high school picks to turn down the draft was
12.99. When they are redrafted, the average pick is round 16.38, a fair drop.