---
title: "About"
author: "Chase Bookin"
date: "February 27, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(ggplot2)
library(janitor)
library(dplyr)
library(gt)
library(readxl)
library(sf)
library(scales)
library(magrittr)
library(haven)
library(infer)
library(Lahman)
library(xml2)
library(rvest)
```
  
  The Major League Baseball draft is an annual event in which the 30 MLB teams
take turns selecting top high school and collegiate prospects. Once drafted, a
player will choose to sign or, if they are a high school senior or college
player with remaining eligibility, not sign and potentially get drafted again
down the road.

  For top high school players especially, the magnitude of this decision is
immense. While certain young players will sign for small contracts, eager to
begin their professional career, others will forgo multi-million dollar signing
bonuses and attend college. For some, the decision pays off handsomely and the
prospect signs for more money as a polished college star. For others, they get
injured or fail to perform in college and never play professional baseball.
Money is obviously not the only factor in a high school prospect's decision -
the team, prospect's health, and family influences among other things can play a
significant role. However, given the amount of money and the opportunity many
high school players opt to delay, it begs the question of how the decision plays
out for these athletes.
  
Do high school players who sign out of high school have a better chance of
making it to the Major Leagues relative to their peers who attend college? What
is the relationship between a player's decision of whether to sign and the
longevity and production of their professional baseball career? Of drafted high
school players that make it to the big leagues, do players that choose to attend
college develop and experience more fruitful careers than players that sign
immediately out of high school? POSSIBLE RESEARCH proportion of picks on high
school players for each MLB team and what proportion make it to the bigs to see
if certain teams better at developing hs talent
  
  The Major League Baseball draft is full of fascinating topics to analyze, and in
this project I will use data to quantify high school players' career outcomes in
relation to this decision.


```{r}
war <- read_csv("test.csv")
```



```{r, include=FALSE}
csv <- read_csv("war.csv")
ortiz <- csv %>%
  filter(name_common == "David Ortiz")
```



```{r, echo=FALSE}

ggplot(data = ortiz, mapping = aes(age, WAR, color = team_ID)) +
  geom_point() +
  theme_classic() + 
  labs(
    x = "Age",
    y = "Wins Above Replacement",
    title = "David Ortiz Wins Above Replacement by Age",
    subtitle = "Note his jump in WAR upon joining the Red Sox"
  )
  
```



  Data: So far, I've loaded in a csv from BaseballReference containing WAR data
for all players which essentially measures their production in the major
leagues. It will be a good source for analyaing the final question I posed
above. My GitHub repo is https://github.com/ChaseBook/final-project.git.

  I have a handful of ideas for how I will begin to play with the data. For one, I
will compare among players that make it to the major leagues, the average
production (WAR) of players who turned down the draft out of high school and
those who signed immediately by position. In addition, I will look at the
average change in draft position for high school players who turn down the
draft. Another idea I will examine is the average speed to the majors in terms
of years spent in the minor leagues or age. Finally, I will look at how high up
organizationally players in each of these two categories are after different
numbers of years in the pros or age. I will also look at career earnings for the
two categories by position to see if the decision pays off on average strictly
from a financial standpoint.











